A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-07-14 paper/fascd.tex
tighten appendix, address a PEF, remove commented-out clutter

%\newcommand{\XX}{{\sffamily X}}
\newcommand{\XX}{\ding{55}}

@@

\begin{abstract}
We present the \emph{full approximation scheme constraint decomposition} (FAScd) multilevel method for solving variational inequalities (VIs).  FAScd is a common extension of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.  We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.  Then, by applying strong smoothers, with work proportional to the number of unknowns on a given mesh level, we demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers.  The example problems include unilateral and bilateral VI problems, with differential operators which are symmetric linear, nonsymmetric linear, and nonlinear.  Our examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
\end{abstract}

%\pef{I want to say it's better than monotone multigrid. Is this too subtle/too much?}
% I vote for not saying this in the abstract.  It is hard to say which aspect of FAScd is more important to beating monotone multigrid: stronger smoother or improved up-smooth CD.  Probably the former, but for now we don't have any illuminating analysis.

@@

%\tableofcontents

\thispagestyle{empty}
%\bigskip

@@

%\pef{I think we need some text here hitting the nail on the head about what's different to monotone multigrid methods (even if it's just the absence of the invasive modifications of the basis functions).}
% Gave it a shot.

@@

\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive Hausdorff measure.  Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.  Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.  For $u,v \in \cV$, $\eps>0$, and $\phi \in \cV'$, define
\begin{equation}
\ip{f(u)}{v} = \eps \left(\grad u, \grad v\right)_{L^2(\Omega)} + b(u,v) - \ip{\phi}{v}. \label{eq:advectiondiffusion}
\end{equation}
%\pef{Why is $\phi \in L^\infty(\Omega)$? If that is the case, then should we have an $L^2$ inner product in the last term, rather than a dual pairing?}
% Fixed.  \phi can be in the dual space
With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$. The interior condition, i.e.~PDE strong form, is the following linear advection-diffusion equation
\begin{equation}
-\eps \grad^2 u + \bX\cdot \grad u = \phi.
\label{eq:advectiondiffusionstrong}
\end{equation}
On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, thus that $f:\cK \to \cV'$ is continuous.  Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.  By the outward flow assumption and the Poincar\'e inequality,
\begin{align*}
\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\dx + b(u-v,u-v) \\
                    &= \eps \int_\Omega |\grad u - \grad v|^2\dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
\end{align*}
Thus $f$ is 2-coercive, and so the associated VI problem is well-posed.
\end{example}

@@

\begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is a given bedrock elevation.  Let $a \in L^2(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (as snow) minus melt and runoff.  Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the following \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} VI problem: find $s \in \mathcal{K} = \{s\ge b\}$ such that
\begin{equation}
\int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad (v-s) \dx \ge \int_\Omega a (v-s)\dx \quad \text{ for all } v \in \mathcal{K}, \label{eq:siavi}
\end{equation}
where $n>1$ and $\Gamma>0$ are constants determined by the physical properties of ice; $n=3$ is a typical value.  The doubly-nonlinear and doubly-degenerate operator in this case is
\begin{equation}
\ip{f(s)}{v} = \int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad v\dx. \label{eq:sia}
\end{equation}
If the bed is flat ($b=0$) then a power transformation converts the problem into $p$-Laplacian form \eqref{eq:plaplacian} with $p=n+1$.  For general beds $b$ less is understood about this operator, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.  This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary from the $s>b$ side.
\end{example}
%\pef{`suggests' is quite weak; can we use a stronger word?}
% Agreed.  Last sentences made brief and direct.

Numerical results are shown in Section \ref{sec:results} for applications of FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
%\pef{I'm not sure starting with the example problems we want to solve is the right move. Let me read more and think about it.}
% I think it is o.k.

@@

%\pef{Can you expand a little bit here? I confess I don't follow.}
% Tried

@@

To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.  Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
\begin{equation}
\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
\end{equation}
\pef{Shouldn't there be some relation between the continuous bounds and the discrete ones?}
In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
\begin{equation}
\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on vertices of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
\end{equation}
the finest-level \emph{constraint set} of admissible FE functions. This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
%As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such vertices.

@@

%\pef{I don't like the name ``monotone restrictions''. The word restriction is strongly associated with operating on the dual space. I would prefer to call them ``monotone injections'', since that is the matching operator.}
% Agreed.  "monotone restriction" is used in GraeserKornhuber2009, but happy not to follow.

@@

\begin{example}  \label{ex:directRbad}

Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).  Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the maximum of $\underline{\gamma}^1$.  Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.  Ignoring Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.  However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematic.  Suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.  In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level: $Pw^0 \notin \mathcal{K}^1$.  On the other hand, suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.  However, while these facts are true, the difficulty with this choice is that the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
\end{example}

@@

Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
\begin{equation}
\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}. \label{eq:fe:chilevels}
\end{equation}
Observe that
\begin{equation}
\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
\end{equation}
A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.  The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.

@@

This construction of multilevel box constraints, fundamentally the same as the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.  Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.

We now compute the LDC differences \cite{GraeserKornhuber2009}:
\pef{Say a few words of motivation first for introducing this concept.}
\begin{equation}
\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}.  \label{eq:fe:philevels}
\end{equation}
We take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.  While $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, they are not ordered as in \eqref{eq:fe:chiordering}.

@@

\begin{pseudofloat}[ht]
\begin{pseudo}
\pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
    $\underline{\chi}^J, \, \overline{\chi}^J = \underline{\gamma}^J - w^J, \, \overline{\gamma}^J - w^J {\large \strut}$ \label{line:vcyclegenchifinest} \\
    for $j=J$ downto $j=1$ \\+
      $\underline{\chi}^{j-1}, \, \overline{\chi}^{j-1} = \maxR \underline{\chi}^j, \, \minR \overline{\chi}^j {\large \strut}$ \label{line:vcyclegenchi} \\
      $\underline{\phi}^j, \, \overline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1}, \, \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \label{line:vcyclegenphi} \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
      $w^{j-1} = \iR(w^j + y^j)$ \label{line:vcyclerestrictsolution} \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \label{line:vcyclerestrictell} \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
    return $w^J+z^J$
\end{pseudo}
\caption{The FAScd V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
\label{alg:fascd}
\end{pseudofloat}

@@

%Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.  Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.  If $m_D$ is the number of vertices $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.  Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.  Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
%\begin{align}
%\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
%I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
%\end{align}
%for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
%\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$). Maybe it would be easier to just use the spelled-out version below, and not use this notation?}
% Agreed.  Greatly simplified

@@

%\pef{I think this is all fairly standard and we could abbreviate this section substantially.}
% Yes.

@@

However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.  The well-known principle of such a cycle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.  Then an admissible initial iterate on the next-finer mesh is generated via prolongation, and so on to finer levels.  (In the VI context one must also truncate after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.
%\pef{Clarify truncation}
% Attempted.

@@

This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.  Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}  Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
%\pef{Explain why?}
% Mostly not done.  But restated for clarity?

@@

%\pef{unlike \cite{GraeserKornhuber2009}?}
% Hard to say.  They only look at one resolution, and they don't show something like "ball".  But their iterations are in the 40--80 range; I've noted that.

>>>>

\newcommand{\XX}{\ding{55}}

@@

\begin{abstract}
We present the \emph{full approximation scheme constraint decomposition} (FAScd) multilevel method for solving variational inequalities (VIs).  FAScd is a common extension both of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.  We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.  Then, by applying strong smoothers, with work proportional to the number of unknowns on a given mesh level, we demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers.  The example problems include unilateral and bilateral VI problems, with differential operators which are symmetric linear, nonsymmetric linear, and nonlinear.  Our examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
\end{abstract}

@@

\thispagestyle{empty}

@@

@@

\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive Hausdorff measure.  Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.  Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.  For $u,v \in \cV$, $\eps>0$, and $\phi \in \cV'$, define
\begin{equation}
\ip{f(u)}{v} = \eps \left(\grad u, \grad v\right)_{L^2(\Omega)} + b(u,v) - \ip{\phi}{v}. \label{eq:advectiondiffusion}
\end{equation}
With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$. The interior condition, i.e.~PDE strong form, is the following linear advection-diffusion equation
\begin{equation}
-\eps \grad^2 u + \bX\cdot \grad u = \phi.
\label{eq:advectiondiffusionstrong}
\end{equation}
On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, thus that $f:\cK \to \cV'$ is continuous.  Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.  By the outward flow assumption and the Poincar\'e inequality,
\begin{align*}
\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\dx + b(u-v,u-v) \\
                    &= \eps \int_\Omega |\grad u - \grad v|^2\dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
\end{align*}
Thus $f$ is 2-coercive, and so the associated VI problem is well-posed.
\end{example}

@@

\begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is a given bedrock elevation.  Let $a \in L^2(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (as snow) minus melt and runoff.  Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the following \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} VI problem: find $s \in \mathcal{K} = \{s\ge b\}$ such that
\begin{equation}
\int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad (v-s) \dx \ge \int_\Omega a (v-s)\dx \quad \text{ for all } v \in \mathcal{K}, \label{eq:siavi}
\end{equation}
where $n>1$ and $\Gamma>0$ are constants determined by the physical properties of ice; $n=3$ is a typical value.  The doubly-nonlinear and doubly-degenerate operator in this case is
\begin{equation}
\ip{f(s)}{v} = \int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad v\dx. \label{eq:sia}
\end{equation}
If the bed is flat ($b=0$) then a power transformation converts the problem into $p$-Laplacian form \eqref{eq:plaplacian} with $p=n+1$.  For general beds $b$ less is understood about this operator, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.  This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary from the $s>b$ side.
\end{example}

Numerical results are shown in Section \ref{sec:results} for applications of FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.

@@

@@

To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.  Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
\begin{equation}
\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
\end{equation}
\pef{Shouldn't there be some relation between the continuous bounds and the discrete ones?}
In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
\begin{equation}
\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on vertices of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
\end{equation}
the finest-level \emph{constraint set} of admissible FE functions. This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.

@@

@@

\begin{example}  \label{ex:directRbad}  
Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).  Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the maximum of $\underline{\gamma}^1$.  Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.  Ignoring Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.  However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematic.  Suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.  In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level: $Pw^0 \notin \mathcal{K}^1$.  On the other hand, suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.  However, while these facts are true, the difficulty with this choice is that the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
\end{example}

@@

Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
\begin{equation}
\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}. \label{eq:fe:chilevels}
\end{equation}
Observe that
\begin{equation}
\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
\end{equation}
A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.  The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.  Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.

@@

This construction of multilevel box constraints, as for the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.  However, when coarsening the problem one does not know what the next coarse corrections will be, and we require constraint sets from which any sum of corrections will be admissible.  With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
\begin{equation}
\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}.  \label{eq:fe:philevels}
\end{equation}
We take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.  While $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, they are not ordered as in \eqref{eq:fe:chiordering}.

@@

\begin{pseudofloat}[ht]
\begin{pseudo}
\pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
    $\underline{\chi}^J, \, \overline{\chi}^J = \underline{\gamma}^J - w^J, \, \overline{\gamma}^J - w^J {\large \strut}$ \label{line:vcyclegenchifinest} \\
    for $j=J$ downto $j=1$ \\+
      $\underline{\chi}^{j-1}, \, \overline{\chi}^{j-1} = \maxR \underline{\chi}^j, \, \minR \overline{\chi}^j {\large \strut}$ \label{line:vcyclegenchi} \\
      $\underline{\phi}^j, \, \overline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1}, \, \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \label{line:vcyclegenphi} \\
      $y^j = 0$ \\
      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
      $w^{j-1} = \iR(w^j + y^j)$ \label{line:vcyclerestrictsolution} \\
      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \label{line:vcyclerestrictell} \\-
    $z^0 = 0$ \\
    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
    for $j=1$ to $j=J$ \\+
      $z^j = y^{j} + P z^{j-1}$ \label{line:vcycleupsmoothinitial} \\
      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
    return $w^J+z^J$
\end{pseudo}
\caption{The FAScd V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
\label{alg:fascd}
\end{pseudofloat}

@@

@@

@@

However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.  The well-known principle of such a cycle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.  Then an admissible initial iterate on the next-finer mesh is generated via prolongation, and so on to finer levels.  (In the VI context one must also truncate after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.

@@

This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.  Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}  Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.

@@

%REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
%$ python3 drape.py -prob ball -levs 9 -o ball -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
%$ mpiexec -n 16 python3 drape.py -prob spiral -levs 10 -o spiral.pvd -fascd_levels_snes_converged_reason -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
%$ paraview ball|spiral.pvd
%use preset "xray" and invert colors; for spiral.pvd, show lgap = u - lb as black if above 0.02; for ball i used 0.001
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
\end{center}
\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
\label{fig:results:classical}
\end{figure}
